\chapter{Overview}\label{overview}

CalChart Version 3.5 is a update of of CalChart Version 3 (the version that has been used by the Cal Band from 1991 to the 2008).  The major improvements have been to port it to the MacOS X platform, integrating closer with the native windowing system functionality like file history and printing. It retains backward compatility with CalChart Version 3 files, though older CalChart version are not supported.

CalChart, like all software, may contain issues and bugs.  And like all software will greatly benefit from feedback from users.  Please document any bugs that you encounter on the CalChart developer website at:

\begin{verbatim}
https://github.com/calband/calchart/issues
\end{verbatim}

Also, we would very much like to hear any suggestions you may have to make CalChart even more useful.

This chapter gives some hints on getting to know Version 3.5.  Much is very similar to Version 3, and for users already familiar with Version 3 may not even notice a difference.

\section{Opening/Creating a Show}\label{newshow}

Starting the program causes a small window to appear.  From the File pull-down menu you may open an existing show via the "Open..." item (or the Alt-O hotkey), or you may create a new show via the "New" item (or the Alt-N hotkey).  You may also choose from the list of files to quickly load a previous open filed.

Creating a new show will start the "New Show Setup Wizard", which is intended to make creating your new show easier.  The wizard can be used to determine the number of points, the kind of labels, and the kind of background.  You must press finish to have the changes take effect, but you choose cancel and then modify any of these settings later on.

\section{Basics}\label{basics}

At the heart of CalChart is the field view.

To scroll the stuntsheet viewer, just move the mouse around the field display area of the window while holding the left mouse button and the control key; scrolling happens automatically.

To zoom in or out of the field view you can use the pinch gesture, the zoom controller, or the control-+ or control-- shortcut.

Use the Undo feature under the Edit pull-down menu to undo actions.  Repeating this causes older and older actions to be undone.  The user can retrace his steps all the way back to the most recent save (providing that the undo buffer, which defaults to 100 items, is big enough).

Various operations can be performed by either clicking on the toolbar icon just above the stuntsheet display area or by using a pull-down menu.  Some features have keystroke commands also.

Use the icons just above the stuntsheet display area to navigate through the show.

\section{Moving Points}\label{movement}

There are several move methods of moving points around the field.  You select different move methods by clicking their toolbar icon.  After a move has been completed, the move method will default back to Translate Points.

\subsection{Selecting Points}\label{selectingpoints}

Points are selected by left-drag the cursor with the mouse press down.  You can choose 3 types of selections: box, clicking to make a polynomial, and a lasso to drag around the points.  Once points are selected many features can be applied to them. Hold down Shift when selecting to add points to an already selected block. Hold down Alt to toggle a points selection.

\subsection{Swap Points}\label{swappoints}

This movement command is used to move swap two different points with each other.

\subsection{Translate Points}\label{translatepoints}

This movement command is used to move individuals or blocks to a new location on the field.

To individual move, left-drag on a point, then move the cursor to the destination, then release the left button.

To block move, first select the block, then hold shift and move any one point of the block as if it were an individual move.

\subsection{Move Points Into a Line}\label{movepointsintoaline}

This movement command is used to move a block into a straight line.

First select the block of points you wish to have placed in a line.  Then left-click the point where the first point in the line should start.  While holding down the left-button, drag the mouse to where the second point in the line should be.  Release the mouse button and CalChart will move all of the points into a straight line from that point.

\subsection{Move Points Into a Shape}\label{movepointsintoashape}

These movement command is used to move a block into several shapes, which are X, cross, box or ellipsis.

First select the block of points you wish to have placed in a shape.  Then left-click the point where the first point in the shape should start.  While holding down the left-button, drag the mouse to where the second extent of the shape.  Release the mouse button and CalChart will move all of the points into a straight line from that point.

\subsection{Move Points Into a Curve}\label{movepointsintoacurve}

These movement command is used to move a block into a curve you drawn on the field.

First select the block of points you wish to have placed in a shape.  Then left-click the point where the first point of the curve.  While holding down the left-button, drag the mouse in the curve that you wish to move the dots.  Release the mouse button and CalChart will move all of the points evenly spaced on the curve.

\subsection{Project Points On a Line}\label{projectpointsonaline}

This movement command is used to move a block onto a projected line.  This is similar to moving dots on a line but the points will extend it the distance, with the space between the first click and the the mouse location dictating the distance between each point.

\subsection{Rotate Block}\label{rotateblock}

This movement command is used to move rotate a block around a point.

First select the block of points you wish to have rotated.  Left-click the pivot point where the block should be rotated around.  On the second left-click a guiding circle should appear.  While holding the left-button, drag around the circle to choose the desired angle.  Release the mouse button and CalChart will rotate the block around the pivot point by the desired angle.

\subsection{Shear Block}\label{shearblock}

This movement command is used to shear a block across around a point.

First select the block of points you wish to have sheared.  Left-click the pivot point where the block should be sheared around.  On the second left-click a guiding line should appear.  The line must extented at a 90 degree angle from the start point to the pivot point.   While holding the left-button, drag the line to the desired shearing amount.  Release the mouse button and CalChart will shear the block around the pivot point by the desired amount.

\subsection{Reflect Block}\label{reflectblock}

This movement command is used to reflect a block around a line.

First select the block of points you wish to have reflected.  Left-click and drag a reflection line.  Release the mouse button and CalChart will reflect all the points their mirror location on the other side of the line.

\subsection{Resize Block}\label{resizeblock}

This movement command is used to resize a block, to make it larger or smaller.

First select the block of points you wish to have resized.  Left-click the central point where the block should be resized around.  Left-click and drag a resizing line.  Release the mouse button and CalChart will resize the block around the central point by the desired amount.

\subsection{Genius Move}\label{geniusmove}

This movement command is used to translate a block of points by describing the general shape you wish the block to attend.

First select the block of points you wish to have moved.  Left-click and drag 3 lines that generally describe the shape you wish the block to take.  Release the mouse button and CalChart will check to see if it is a valid move, and if so, will apply a genious move to the block.

\section{Labels and Symbols}\label{symbols}

Select a set of points, then click on one of the icons just above the stuntsheet display area for the desired action.

\section{Field Properties}\label{fieldproperties}

There are a variety of controls that change the properties of the field.

\subsection{Grid Spacing}\label{gridspacing}

The Grid Spacing controls the distance between where dots may be placed on the field.

\subsection{Zoom}\label{zoomlevel}

You can adjust the magnification of the field with the zoom level, or set to fit to see the whole field at once.

\subsection{Reference Group}\label{referencegroup}

Each point has 3 reference points to assist with more complicated animation commands.  You can set which reference point is active by changing the Reference group.

\subsection{Draw Paths}\label{drawpaths}

The Draw Paths option causes a preview of the path that point will take to be displayed. 

\subsection{Ghost Sheet}\label{ghostsheet}

A Ghost sheet is the image of the a different sheet semi-imposed on the current sheet.
You can choose to display the next, previous, or a specific page.

\section{Editing Continuity}\label{editcont}

The Continuity commands are displayed on the right side of the field.  By double clicking on an entry, you can edit the command in place.

Also, you can Open an Edit Continuity window (under the Edit pull-down menu) to type
continuity commands.

\section{Printed Continuity}\label{printcont}

This is not yet finished but it mostly works for viewing the continuity as it will be printed.  Open a Printed Continuity window under the Edit pull-down menu.

\section{Animation}\label{animwin}

Open an Animation window (under the Window pull-down menu) to view animation.  Use the icons to navigate through the show and the slider to set the tempo.  Click on the check-box labeled Collisions to enable detection of collisions.  When errors are encountered while generating the animation, a window will appear with the types of errors.  You can
click on an error to select the points that caused this error.

