\documentstyle[a4,texhelp]{report}
\title{CalChart User's Manual 3.2.2}
\author{Richard Powell}
\date{June 21, 2011}
\makeindex
\begin{document}
\maketitle
\tableofcontents

\chapter*{Copyright Notice}\label{copyright}

Copyright (c) 1994-2011  Garrick Brian Meeker, Richard Michael Powell

This program is free software: you can redistribute it and/or modify
it under the terms of the GNU General Public License as published by
the Free Software Foundation, either version 3 of the License, or
(at your option) any later version.

This program is distributed in the hope that it will be useful,
but WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
GNU General Public License for more details.

\chapter*{Acknowledgments}\label{acknowledge}

Acknowledgments for Version 3.2.2

This manual was updated by Richard Powell.

Many thanks to Jonathan Nagata (DM '10) and Allan Yu (DM '11) for helping
with design and testing of the new version updated since 2010.

Also thanks to those developers how have been involved since 2010:
David Clarino, Brian Wolf, John Wang, and of course, Garrick Meeker.

Acknowledgments for Version 3.0

Portions of this manual were written by Daniel Isaksen and Martin Haye.

Many thanks to Dan Isaksen for assisting in the design and testing of
this program.

Thanks to Rommel Suratos (DM '94) for supporting the initial PostScript
printing routines that were the beginnings of this program and to Rob
McGuire (DM '98) and his Stunt committee for supporting this first full
use of this program.

And of course, thanks to Martin Haye for writing the initial version,
and to Greg Jones (DM '90) for supporting the concept of CalChart.

\chapter{Introduction}

CalChart is a tool for designing shows for the University of California
Marching Band.

\input{gpl.tex}
\input{install.tex}
\input{overview.tex}
\input{refer.tex}
\input{bugs.tex}
\input{anim.tex}

\chapter{Future of CalChart}\label{future}

Calchart is under continuous development.  Follow the activity at our
project page at:
\begin{verbatim}
http://sourceforge.net/projects/calchart
\end{verbatim}


\end{document}
